Transactional Memory Concurrency Verification with Landslide

### 2018-03-12 htm/paper.tex
state first thm

@@

Left unstated in interleavings such as $\ti@1-2c, \tj@1-3, \ti@4-7$%
\footnote{For the sake of example I reorder \ti's write to {\tt x} before \tj's part here.}
are HTM's execution semantics, namely:
\begin{enumerate}
	\item any modifications to shared state (such as $2c$) by \ti are not visible to \tj during its execution,
		despite \tj being executed afterwards, and
	\item all local and global state changes by \ti between lines $1$ and $2c$ are discarded when jumping to line $4$.
\end{enumerate}

@@

@@

{\bf Transaction failure codes.}
When a transaction fails, {\tt \_xbegin()} returns a failure code
denoting the reason (or combination of reasons) therefor \cite{htm-gcc}.
If a program then cases on that failure code to select between different backup paths,
model checking it by simply injecting a single type of failure may be unsound.
For example, a program executing a transaction which is guaranteed to never conflict with any other threads,
and hence never abort without {\tt \_XABORT\_RETRY},
could legally {\tt assert(false);} in its failure path,
while our approach in Section~\ref{sec:design} would erroneously trigger that assertion and report a bug.
Likewise, the transactional data structures from \cite{htm-mario}
abstract away any spurious {\tt \_XABORT\_RETRY} aborts behind a retry loop;
a MC unwise to that idiom would call it an infinite loop bug.

>>>>

This section introduces the fundamental concepts and prior work
in both hardware transactional memory and stateless model checking,
which we propose to combine.

@@

Left unstated in interleavings such as $\ti@1-2c, \tj@1-3, \ti@4-7$%
\footnote{For the sake of example I reorder \ti's write to {\tt x} before \tj's part here.}
are HTM's execution semantics, namely:
\begin{enumerate}
	\item any modifications to shared state (such as $2c$) by \ti are not visible to \tj during its execution,
		despite \tj being executed afterwards, and
	\item all local and global state changes by \ti between lines $1$ and $2c$ are discarded when jumping to line $4$.
\end{enumerate}
% there shouldbe a better word than 'requires' here.. benefits from.. "the point of it is"..
While use of TSX in production requires the performance advantage
of temporarily staging such accesses in local CPU cache,
model checking such programs need be concerned only with the program's {\em observable} behaviours.
We claim that MCing the interleaving $\ti@1, \tj@1-3, \ti@4-7$
is an equivalent verification as MCing the one above;
in fact, this interleaving suffices to check
all observable behaviours
of all interleavings
of all subsets of $\tj@1-3$
with all subsets of $\ti@2a-2c$,
whether they share a memory conflict or not.
Stated formally:

\renewcommand\proofname{Proof Sketch}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\newcommand\tii{\ensuremath{\hilight{lavender}{\mathbf{Ti}}}\xspace}
\newcommand\tjj{\ensuremath{\hilight{seafoam}{\mathbf{Tj}}}\xspace}
\newcommand\tkk{\ensuremath{\hilight{salmon}{\mathbf{Tk}}}\xspace}

\begin{theorem}[Equivalence of Aborts]
	\label{thm:equiv}
	Let:%
	\footnote{Arbitrary code may not be structured to distinguish these as nicely as in our examples;
	e.g., more code may exist in the success branch after {\tt \_xend()};
	such would be considered part of $\omega$ here.}
	\begin{itemize}
		\item $\tii@\alpha$ be an HTM begin operation,
		\item $\tii@\beta_1\dots\tii@\beta_n$ be the transaction body (with $\beta_n$ the HTM end call),
		\item $\tii@\phi_1\dots\tii@\phi_m$ be the failure path, and
		\item $\tii@\omega_1\dots\tii@\omega_l$ be the subsequent code executed unconditionally.
	\end{itemize}
	Then, for any interleaving prefix
	\[
		\tii@\alpha,\tii@\beta_1\dots\tii@\beta_b,\tjj@\gamma_1\dots\tjj@\gamma_j,\tkk@\kappa_1\dots\tkk@\kappa_k
	\]
	with $b<n$, $j \ne i$, $k \ne i$, etc., either:
	\begin{enumerate}
		\item $\tii@\alpha,\tii@\beta_1\dots\tii@\beta_b\dots\tii@\beta_n,\tjj@\gamma_1\dots\tjj@\gamma_j,$ \\
			$\tkk@\kappa_1\dots\tkk@\kappa_k$
			(independent case), or
		\item $\tii@\alpha,\tjj@\gamma_1\dots\tjj@\gamma_j,\tkk@\kappa_1\dots\tkk@\kappa_k,\tii@\phi_1\dots$
			(conflicting case)
	\end{enumerate}
	is observationally equivalent.
\end{theorem}

\begin{proof}
	asdfasdf
\end{proof}

@@

Next, we address the memory accesses within transactions with regard to data-race analysis.
From Theorem~\ref{thm:equiv} we have that the body of all transactions may be executed atomically within the MC environment.
While they may interleave within other non-transactional sequences,
no other operations (whether transactional or not) will interrupt them.
We claim this level of atomicity is equivalent to that provided by a global lock,
and hence abstracting it as such in Landslide's data-race analysis is sound.

\begin{theorem}[Transactions are a Global Lock]
	asdfasdfaf
\end{theorem}

\begin{proof}
	asdfasdf
\end{proof}

@@

This section discusses opportunities for future work.

{\bf Transaction failure codes.}
When a transaction fails, {\tt \_xbegin()} returns a failure code
denoting the reason, or combination thereof, therefor \cite{htm-gcc}.
If a program then cases on that failure code to select between different backup paths,
model checking it by simply injecting a single type of failure may be unsound.
For example, a program executing a transaction which is guaranteed to never conflict with any other threads,
and hence never abort without {\tt \_XABORT\_RETRY},
could legally {\tt assert(false);} in its failure path,
while our approach in Section~\ref{sec:design} would erroneously trigger that assertion and report a bug.
Likewise, the transactional data structures from \cite{htm-mario}
abstract away any spurious {\tt \_XABORT\_RETRY} aborts behind a retry loop;
a MC unwise to that idiom would call it an infinite loop bug.
